---
title: "Brown Trout stocking summary (GLFC)"
author: "Colin Lake - MNRF<br>Glenora Fisheries Station<br>Lake Ontario Management Unit"
date: "Date last updated: `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true

---

<style type="text/css">
table {
  font-size: 9pt;
}
</style>

<style type="text/css">
p.caption {
  font-size: 1.5em;
}
</style>

```{r readData,include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE)
```

```{r setup}

#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=8)
source("~/projects/themes/theme_gfs.R")

library(tidyr)
library(dplyr)
library(readxl)
library(DT)
library(ggsci)
library(sjmisc)
library(janitor)
library(plotly)
library(lkonstock)
library(wesanderson)
library(crosstalk)
library(leaflet)
library(rgdal)
library(lubridate)
library(modelsummary)

df <- read.csv("Data/Raw/bt_glfc.csv") # previously downloaded - brown trout only

# only one state has entered 2020 data, so let's remove that

df <- df %>% subset(year !="2020")

# first, fix up the data set so it's more usable

# bring in lookup table from GLFC data template - make stage variable that is 'plain language' instead of a code
stglkup <- read_xlsx("Data/Lookup/GLFSD_NewDataSubmission_Template.xlsx", sheet = "Lifestage")
stglkup <- stglkup %>% rename(life_stage = stage)
stglkup$stage <- gsub("(.*),.*", "\\1", stglkup$description)
stglkup$stage[stglkup$life_stage == "a"] <- "adult"
stglkup <- subset(stglkup, select=-c(description))
#df <- left_join(df, stglkup, by = c("life_stage" = "stage"))
df <- left_join(df, stglkup)

# there are a small number (n=26) of events with unknown life stage - by age, it looks like they should be yearlings
df$stage[df$life_stage == "u" & df$age_months <=24] <- "yearling"
df$stage[df$life_stage == "u" & df$age_months >24] <- "adult"


# weird variable names - fix them
df <- rename(df,lake = X_lake)
df <- rename(df,clip = X_clip)
df <- rename(df,strain = X_strain)

# concatenate the date values for use later
df$date <- as.Date(with(df, paste(year, month, day,sep="-")), "%Y-%m-%d")

minyr <- min(df$year)
maxyr <- max(df$year)


# yearling equivalents - these values used to calc. 
yearling_mult <- read.csv(textConnection("
stage, yearling_mult
adult, 1
yearling, 1
fingerling, 10
spring fingerling, 10
summer fingerling, 10
fall fingerling, 10
fry, 100
"))


df <- left_join(df, yearling_mult)

df$yreq <- as.integer(df$number_stocked/df$yearling_mult)

```
<br><hr style="border:1px solid gray"> </hr><br>

## Data source
These data (`r minyr`- `r maxyr`) are from the GLFC Stocking Database.  Since the new stocking data-entry and viewing portal is currently being developed, data are only complete up to 2019.  (Michigan is the only exception - they have 2020 data entered, but i omitted it for the following plots, etc).

These are pretty rough visualizations - just meant to inspire some ideas.  There is a lot of work going on to improve the quality of the GLFC stocking data base, but we're not there yet - this data set is still rough in places.

<br><hr style="border:1px solid gray"> </hr><br>

## lifestage
This table shows the age (min and max) of the various Brown Trout life stages, and the time of year (month) that they were stocked. n = number of stocking events of the particular life stage.

```{r lifestage}


chkstg1 <- dplyr::count(df, stage) # <- count stocking events by lifestage

df$stage <- factor(df$stage, 
                          levels = c(
                            "fry",
                            "fingerling",
                            "spring fingerling",
                            "summer fingerling",
                            "fall fingerling",
                            "yearling",
                            "adult")
                       )

chkstg2 <- df %>% group_by(stage) %>% 
  summarise(min_age = min(age_months, na.rm = TRUE),
            max_age = max(age_months, na.rm = TRUE),
            min_month = min(month, na.rm = TRUE),
            max_month = max(month, na.rm = TRUE))  

chkstg <- left_join(chkstg2, chkstg1)

chkstg <- select(chkstg, stage, n, everything())


rm(chkstg1)
rm(chkstg2)

library(kableExtra)
kbl(chkstg, format.args = list(big.mark = ","),row.names = FALSE)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = TRUE))%>%
  row_spec(nrow(chkstg)-0, extra_css = "border-bottom: 1.5px solid")

```


<br><hr style="border:1px solid gray"> </hr><br>

## Number Stocked (yearling eq) {.tabset .tabset-fade .tabset-pills}

All lifestages converted to yearling equivalents

### Bar Plot - Number by Lake
```{r number by lake plot, fig.width=10, fig.height= 7}

library(plotly)
#library(wesanderson)
library(unikn)


# note that we are using yearling equivalents that were calculated earlier

dfnum <- data.frame(
  group_by(df, lake, state_prov, year, stage,) %>%summarise(number = sum(yreq))
  )

stagechk <- dplyr::count(dfnum, vars = stage) # make sure all lifestages are included when factors are ordered

pal1 <- wes_palette("Zissou1",12,type = "continuous")
pal1 <- as.character(pal1)

# all tribs ----
fig1 <- plot_ly(dfnum, x = ~year, y = ~number, color = ~lake, type = 'bar',
                colors = pal1,
                hoverinfo = "text",
                hovertext = paste(
                  "<b>lake: </b>", dfnum$lake, 
                  "<br><b>lifestage: </b>", dfnum$stage, 
                  "<br><b>year: </b>", dfnum$year, 
                  "<br><b>number: </b>", scales::comma(round(dfnum$number), accuracy = 1))
                )

fig1 <- fig1 %>% layout(xaxis = list(title = 'YEAR'), yaxis = list(title = 'NUMBER'), barmode = 'stack')

fig1 <- fig1 %>% layout(
  xaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)), 
  yaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)),
  legend = list(font = list(size = 24))
  )

fig1
```
<br><hr style="border:1px solid gray"> </hr><br>

### Bar Plot - Number by State or Province
```{r number by state plot, fig.width=10, fig.height= 7}

library(plotly)
library(wesanderson)
library(unikn)

dfnum <- data.frame(
  group_by(df, state_prov, year, stage,) %>%summarise(number = sum(yreq))
  )

stagechk <- dplyr::count(dfnum, vars = stage) # make sure all lifestages are included when factors are ordered


# all tribs ----
fig1 <- plot_ly(dfnum, x = ~year, y = ~number, color = ~state_prov, type = 'bar',
                colors = pal1,
                hoverinfo = "text",
                hovertext = paste(
                  "<b>state_prov: </b>", dfnum$state_prov, 
                  "<br><b>lifestage: </b>", dfnum$stage, 
                  "<br><b>year: </b>", dfnum$year, 
                  "<br><b>number: </b>", scales::comma(round(dfnum$number), accuracy = 1))
                )

fig1 <- fig1 %>% layout(xaxis = list(title = 'YEAR'), yaxis = list(title = 'NUMBER'), barmode = 'stack')

fig1 <- fig1 %>% layout(
  xaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)), 
  yaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)),
  legend = list(font = list(size = 24))
  )

fig1
```

<br><hr style="border:1px solid gray"> </hr><br>

## Mean weight of yearlings by Great Lake

Very quick look at what the data look like - obviously some issues.  Check to see if grams entered instead of kilograms in some instances.

```{r average weight}

# FISH WEIGHT

fishweight <- df %>% filter(stage == "yearling")

fishweight$weight <- fishweight$total_weight_kg*1000/fishweight$number_stocked

fishweight <- tbl_df(fishweight) %>%
  group_by(year, lake) %>%
  summarise(N = length(year), avg = round(mean(weight, na.rm=T),2), 
            sd = round(sd(weight, na.rm=T),2), se = round(sd/sqrt(N),2)) 

fig2 <-ggplot(fishweight, aes(x=year, y=avg, label = avg)) + 
  ylab("Weight g") + xlab("Year Stocked") +
  labs (title = "Spring Yearlings - avg. size")+ 
  geom_errorbar(aes(ymin=avg-sd, ymax=avg+sd), width=.5) +
  geom_point(colour="black", shape=19, size = 3) + 
  #geom_hline(aes(yintercept = mean_temp), color="grey", size = 1, linetype = 5) +
  #geom_text(aes(maxyr-1, mean_temp,label = mean_temp, vjust = -0.5), color = "red")+
  theme(legend.title=element_blank())+scale_color_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous (breaks = seq(minyr, maxyr, by=3))+
  theme(legend.position = c(.2, .8)) + 
  theme_gfs()+ 
  theme(axis.text.x=element_text(size=12))+
  theme(axis.text.x=element_text(angle = 90))+
  geom_text(hjust = 0, nudge_x = 0.12, size = 3) + 
  expand_limits(x = maxyr + 0.5)

fishweight$lake <- as.factor(fishweight$lake)

fig2 + facet_grid(lake ~., scales="free") +
  theme(strip.background =element_rect(fill="white"))+
  theme(panel.spacing = unit(2, "lines"))


```

<br><hr style="border:1px solid gray"> </hr><br>


## Stocking Dates
Average dates of Brown Trout stocking events (yearlings only; averaged by lake and year)
```{r stocking_date}

# Check this carefully - surplus brood, etc., can be stocked at unusual times - omit these outliers.
dfdate <- df %>% filter(stage == "yearling")

#omit_stage <- c("Egg", "Adult", "Sub-adult", "Fry")

dfdate <- subset(dfdate, stage == "yearling")

dfdate <- dfdate %>%
  mutate(date = make_date(year, month, day))

dfdate$ordinal <- format(dfdate$date, "%j")
dfdate$ordinal <- as.numeric(dfdate$ordinal)

releasedate <- tbl_df(dfdate) %>%
  group_by(year, lake) %>%
  summarise(N = length(year), avg = round(mean(ordinal, na.rm=T),2), 
            sd = round(sd(ordinal, na.rm=T),2), se = round(sd/sqrt(N),2)) 


# find average release date by life stage

lakeavg <- releasedate %>%
  group_by(lake) %>%
  summarise(lakeavg = as.integer(mean(avg, na.rm=TRUE)))

releasedate <- left_join(releasedate, lakeavg)

# # add these columns so a label can be built
releasedate$mon = 1
releasedate$day = 1
releasedate$start <- as.Date(with(releasedate, paste(year, mon, day, sep = "-")), "%Y-%m-%d")
releasedate$average <- as.Date((releasedate$start) + releasedate$avg)
releasedate$datelabel <- format(releasedate$average, "%b %d")

# vertical plot

fig4 <-ggplot(releasedate, aes(x=year, y=avg, label = datelabel)) + 
  ylab("ordinal Day") + xlab("Year Stocked") +
  #labs (title = "Average Dates of Atlantic Salmon Stocking (life stage/year)")+ 
  geom_errorbar(aes(ymin=avg-sd, ymax=avg+sd), width=.5) +
  geom_point(colour="black", shape=19, size = 3) + 
  geom_hline(data = releasedate, aes(yintercept = lakeavg), color="grey", size = 1, linetype = 5) +
  theme(legend.title=element_blank())+scale_color_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous (breaks = seq(minyr, maxyr, by=3))+
  theme(legend.position = c(.2, .8)) + 
  theme_gfs()+ 
  theme(axis.text.x=element_text(size=12))+
  theme(axis.text.x=element_text(angle = 90))+
  geom_text(hjust = 0, nudge_x = 0.12, size = 3) + 
  expand_limits(x = 2020.5)

fig4 + facet_grid(lake ~., scales="free") +
  theme(strip.background =element_rect(fill="white"))

```

<br><hr style="border:1px solid gray"> </hr><br>

## Map

Stocking events aggregated by stocking site (spring yearlings only).  Radius of circle indicates biomass (kg) stocked

```{r MAP}

stocked <- df %>% subset(
  stage == "yearling"
)

stocked <- stocked%>%
  group_by(
    year,
    lake,
    state_prov,
    location_primary,
    latitude,
    longitude)%>%
  summarise(
    number = sum(yreq),
    biomass = sum(total_weight_kg)
  )



palette <- colorFactor(palette = rainbow(length(unique(stocked$state_prov))), domain = stocked$state_prov)

sd <- SharedData$new(stocked)


map <- leaflet(width = "100%", height = 800) %>%
  addProviderTiles("Esri.WorldStreetMap", group = "Road map") %>%
  addProviderTiles("Esri.WorldImagery", group = "Satellite") %>%
  addProviderTiles("CartoDB.Positron", group = "Gray Scale") %>%
  addCircleMarkers(data = sd, lat = ~latitude, lng = ~longitude, radius = ~biomass/100,
                   stroke = TRUE,
                   color = "black",
                   weight = 1,
                   fillColor = ~palette(stocked$state_prov),
                   fillOpacity = .5,
                   label=~as.character(location_primary),
                   popup = paste
                   ("<b><font size=+1><font color=#5499C7>",stocked$location_primary, "</b></font size></font color><br>",
                     "_______________________<br>",
                     "<br><b>Year: </b>", stocked$year,
                     "<br><b>Number: </b>", stocked$number,
                     "<br><b>Biomass (kg): </b>", stocked$biomass,
                     "<br><b>LAT: </b>", stocked$latitude,
                     "<br><b>LON: </b>", stocked$longitude
                   ))%>%

  addLayersControl(position = 'topright',
                   baseGroups = c("Gray Scale",
                                  "Satellite"))

bscols(widths = c(3,12),
       list(
         filter_select(id = "year", label = "Year", sharedData = sd, group = ~year, multiple = FALSE)
         #filter_slider(id = "STOCKING_YEAR", label = "Year", sharedData = sd, column = ~STOCKING_YEAR, step = 1, sep = "", width = "100%")
         ),
  map
)


```

<br><hr style="border:1px solid gray"> </hr><br>

<a href="#header">back to top</a>