---
title: "Brown Trout stocking summary (LOMU)"
author: "Colin Lake - MNRF<br>Glenora Fisheries Station<br>Lake Ontario Management Unit"
date: "Date last updated: `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true

---

<style type="text/css">
table {
  font-size: 9pt;
}
</style>

<style type="text/css">
p.caption {
  font-size: 1.5em;
}
</style>

```{r readData,include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE)
```

```{r setup}

#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=8)
source("~/projects/themes/theme_gfs.R")

library(tidyr)
library(dplyr)
library(readxl)
library(DT)
library(ggsci)
library(sjmisc)
library(janitor)
library(plotly)
#library(lkonstock) this is a data package I wrote, but I need some help to get it into a shareable format.  For now, i just included the .rda file
library(wesanderson)
library(crosstalk)
library(leaflet)
library(rgdal)

load(file = "Data/Processed/lkonstock.rda")

df <- lkonstock

# choose the range of years to present

minyr <- 2001
maxyr <- 2021

df <- df %>% filter(SPECIES_COMMON_NAME == "Brown Trout" & STOCKING_YEAR >=minyr & lkonstock$STOCKING_YEAR <= maxyr)

```
<br><hr style="border:1px solid gray"> </hr><br>

## Data source
These data are from MNRF's Fish Stocking Information System (FSIS: `r minyr`- `r maxyr`).  This summary is for LOMU Brown Trout stocking events only.  If you see anything that looks like an error, please let me know.

<br><hr style="border:1px solid gray"> </hr><br>

## Lifestage
This table shows the age (min and max) of the various Brown Trout life stages, and the time of year (month) that they were stocked. n = number of stocking events of the particular life stage.

```{r lifestage}

chkstg1 <- dplyr::count(df, LIFESTAGE) # <- CHECK LIFESTAGE

chkstg2 <- df %>% group_by(LIFESTAGE) %>% 
  summarise(min_age = min(FISH_AGE, na.rm = TRUE),
            max_age = max(FISH_AGE, na.rm = TRUE),
            min_month = min(STOCKING_MONTH, na.rm = TRUE),
            max_month = max(STOCKING_MONTH, na.rm = TRUE))  

chkstg <- left_join(chkstg1, chkstg2)

chkstg <- chkstg[order(chkstg$min_age),]


rm(chkstg1)
rm(chkstg2)

library(kableExtra)
kbl(chkstg, format.args = list(big.mark = ","),row.names = FALSE)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = TRUE))%>%
  row_spec(nrow(chkstg)-0, extra_css = "border-bottom: 1.5px solid")

```


<br><hr style="border:1px solid gray"> </hr><br>

## Number Stocked {.tabset .tabset-fade .tabset-pills}

### Bar Plot
```{r number plot, fig.width=10, fig.height= 7}

library(plotly)
library(wesanderson)
library(unikn)

dfnum <- data.frame(
  group_by(df, STOCKING_YEAR, LIFESTAGE) %>%summarise(NUMBER = sum(FISH_STOCKED_COUNT))
  )

stagechk <- dplyr::count(dfnum, vars = LIFESTAGE) # make sure all lifestages are included when factors are ordered

dfnum$LIFESTAGE <- factor(dfnum$LIFESTAGE, 
                          levels = c(
                            "Spring Fingerling",
                            "Fall Fingerling",
                            "Spring Yearling", 
                            "Adult"))


# all tribs ----
fig1 <- plot_ly(dfnum, x = ~STOCKING_YEAR, y = ~NUMBER, color = ~LIFESTAGE, type = 'bar',
                colors = (pal_jama("default")(4)),
                hoverinfo = "text",
                hovertext = paste(
                  "<b>Lifestage: </b>", dfnum$LIFESTAGE, 
                  "<br><b>Year: </b>", dfnum$STOCKING_YEAR, 
                  "<br><b>Number: </b>", scales::comma(round(dfnum$NUMBER), accuracy = 1))
                )

fig1 <- fig1 %>% layout(xaxis = list(title = 'YEAR'), yaxis = list(title = 'NUMBER'), barmode = 'stack')

fig1 <- fig1 %>% layout(
  xaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)), 
  yaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)),
  legend = list(font = list(size = 24))
  )

fig1
```
<br><hr style="border:1px solid gray"> </hr><br>


### Table
MNRF-stocked fish are almost exclusively Spring Yearlings.  Other life stages are generally stocked by community hatcheries. 

```{r numbers}

# dfnum <- data.frame(
#   group_by(df, STOCKING_YEAR, LIFESTAGE) %>%summarise(NUMBER = sum(FISH_STOCKED_COUNT))
#   )

options(knitr.kable.NA = '--')


btnum <- spread(dfnum, LIFESTAGE, NUMBER)
btnum <- select(btnum, 
                "STOCKING_YEAR",
                "Spring Fingerling",
                "Fall Fingerling",
                "Spring Yearling")
btnum <- adorn_totals(btnum, where = "row")
#btnum <- btnum %>% replace(is.na(.), 0) %>% mutate(Total = rowSums(.[3:8]))
btnum <- btnum %>% mutate(Total = rowSums(.[2:4], na.rm = TRUE))
btnum <- btnum %>%
  rename(
    YEAR = STOCKING_YEAR
  )

library(kableExtra)
kbl(btnum,format.args = list(big.mark = ","))%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = TRUE))%>%
  row_spec(nrow(btnum)-1, extra_css = "border-bottom: 1.5px solid")

```

<br><hr style="border:1px solid gray"> </hr><br>

<br><hr style="border:1px solid gray"> </hr><br>

## Number by Stocking Zone {.tabset .tabset-fade .tabset-pills}
Total number of Brown Trout stocked by stocking zone (most recent five years; 2001 - current).  The map at the bottom of this page shows the location of the various stocking zones.  
<br>

### Recent five years
```{r 5 yr stocking zones}

stocksum <- df%>%
  group_by(
    STOCKING_YEAR, 
    STOCKING_ZONE)%>%
  summarise(NUMBER = sum(FISH_STOCKED_COUNT)
  )

stocksum5 <- stocksum %>%
  filter(STOCKING_YEAR > maxyr-5)

# need to make complete data frame with all missing values so plotly will work properly
STOCKING_YEAR <- rep((c(min(stocksum5$STOCKING_YEAR)):(max(stocksum5$STOCKING_YEAR))), times=7) # all years, repeated 7 times (once per stocking zone)
STOCKING_ZONE <- rep(c(1:7), times=5)
dfx <- as.data.frame(cbind(STOCKING_YEAR,STOCKING_ZONE))
dfx$NUMBER <- NA

dfy <- left_join(dfx, stocksum5, by =c("STOCKING_YEAR", "STOCKING_ZONE"))%>%
  mutate(NUMBER = coalesce(NUMBER.x, NUMBER.y)) %>%
  select(-NUMBER.x, -NUMBER.y)

dfy$NUMBER[is.na(dfy$NUMBER)] <- 0

# rename back to original dataframe
stocksum5 <- dfy

# https://stackoverflow.com/questions/39961798/how-to-complete-missing-factor-levels-in-data-frame
#stocksum5 <- complete(stocksum5, STOCKING_YEAR, STOCKING_ZONE, fill = list(NUMBER=0))

# https://stackoverflow.com/questions/42858202/r-plotly-and-subplot-fastest-way-to-create-a-subplot-based-on-a-factor

stocksum5$STOCKING_ZONE <- paste0("Zone ", stocksum5$STOCKING_ZONE)

splitted_list <- split(stocksum5, stocksum5$STOCKING_ZONE)

plot_list <- lapply(splitted_list, plot_ly, x = ~STOCKING_YEAR, y = ~NUMBER, type = "bar",  name = ~STOCKING_ZONE, yaxis2 = list(showticklabels = F))  

# legend=list(title=list(text='<b> Zone </b>'))

subplot(plot_list, margin = .01, shareY = TRUE) 



```

### 2001 - current

```{r all years stocking zones}

stocksum <- df%>%
  group_by(
    STOCKING_YEAR, 
    STOCKING_ZONE, 
    SPECIES_COMMON_NAME)%>%
  summarise(NUMBER = sum(FISH_STOCKED_COUNT)
  )



# https://stackoverflow.com/questions/42858202/r-plotly-and-subplot-fastest-way-to-create-a-subplot-based-on-a-factor

stocksum$STOCKING_ZONE <- paste0("Zone ", stocksum$STOCKING_ZONE)

splitted_list <- split(stocksum, stocksum$STOCKING_ZONE)

plot_list <- lapply(splitted_list, plot_ly, x = ~STOCKING_YEAR, y = ~NUMBER, type = "bar",  name = ~STOCKING_ZONE, yaxis2 = list(showticklabels = F))

# legend=list(title=list(text='<b> Zone </b>'))

subplot(plot_list, margin = .01, shareY = TRUE)



```

<br><hr style="border:1px solid gray"> </hr><br>

## Mean Weight by Lifestage

MNRF-produced fish only; click on the legend to isolate life stages.

```{r average weight}

# FISH WEIGHT
#  https://plotly.com/r/custom-buttons/#update-button use update buttons for MNRF vs CHP vs TOTAL

dfmnr <- subset(df, df$PROPONENT_TYPE_CODE == "MNR")


fishweight <- tbl_df(dfmnr) %>%
  group_by(STOCKING_YEAR, LIFESTAGE) %>%
  summarise(N = length(STOCKING_YEAR), avg = round(mean(FISH_WEIGHT, na.rm=T),2), 
            sd = round(sd(FISH_WEIGHT, na.rm=T),2), se = round(sd/sqrt(N),2)) 

omit_stage <- c("Egg", "Adult", "Sub-adult", "Fry")
fishweight <- subset(fishweight, !(LIFESTAGE %in% omit_stage))

fishweight$LIFESTAGE <- factor(fishweight$LIFESTAGE, levels = c(levels = c("Spring Fingerling", "Fall Fingerling", "Spring Yearling")))

fig2 <- plot_ly(data = fishweight[which(fishweight$LIFESTAGE == 'Spring Fingerling'),], x = ~STOCKING_YEAR, y = ~avg, type = 'scatter', mode = 'markers',
               name = 'Spring Fingerling',
               marker = list(size = 11), 
               error_y = ~list(array = sd,
                               color = '#000000',
                               width = 8,
                               thickness = 1),
               hovertemplate = paste('<b>Year: </b> %{x} <br><b>Weight: </b>%{y} <extra></extra>'))

fig2 <- fig2 %>% add_trace(data = fishweight[which(fishweight$LIFESTAGE == 'Fall Fingerling'),], name = 'Fall Fingerling')
fig2 <- fig2 %>% add_trace(data = fishweight[which(fishweight$LIFESTAGE == 'Spring Yearling'),], name = 'Spring Yearling')

fig2 <- fig2 %>% layout(xaxis = list(title = 'YEAR'), yaxis = list(title = 'MEAN WT. (g)'))
fig2 <- fig2 %>% layout(xaxis = list(tickfont = list(size = 15)), yaxis = list(tickfont = list(size = 15)))

fig2


```

<br><hr style="border:1px solid gray"> </hr><br>

## Biomass {.tabset .tabset-fade .tabset-pills}

### Bar plot

```{r biomass plot}

library(plotly)

dfmass <- df

dfmass <- filter(dfmass, dfmass$LIFESTAGE != "Egg")

dfmass <- dfmass %>%
  group_by(
    LIFESTAGE,
    STOCKING_YEAR,
  )

dfmass <- dfmass %>% summarise(
  BIOMASS = round(sum(BIOMASS_KG),digits=0)
)

dfmass$LIFESTAGE <- factor(dfmass$LIFESTAGE, levels = c("Fry", "Spring Fingerling", "Fall Fingerling", "Spring Yearling", "Sub-adult", "Adult"))

fig3 <- plot_ly(dfmass, x = ~STOCKING_YEAR, y = ~BIOMASS, color = ~LIFESTAGE, type = 'bar',
                colors = (pal_jama("default")(4)),
                hoverinfo = "text",
                hovertext = paste(
                  "<b>Lifestage: </b>", dfmass$LIFESTAGE, 
                  "<br><b>Year: </b>", dfmass$STOCKING_YEAR, 
                  "<br><b>Biomass: </b>", dfmass$BIOMASS))

fig3 <- fig3 %>% layout(xaxis = list(title = 'YEAR'), yaxis = list(title = 'BIOMASS (kg)'),barmode = 'stack')

fig3 <- fig3 %>% layout(
  xaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)), 
  yaxis = list(tickfont = list(size = 20), titlefont = list(size = 24)),
  legend = list(font = list(size = 24))
  )


fig3

```


<br><hr style="border:1px solid gray"> </hr><br>

### Table
Note: biomass is given in kg.

```{r biomass}

btwt <- data.frame(
    group_by(df, STOCKING_YEAR, LIFESTAGE) %>%summarise(BIOMASS = sum(BIOMASS_KG)))


# adults are one-time deal; omit them

btwt <- btwt%>%
  filter(LIFESTAGE != "Adult")

#library(tidyr)
#library(janitor)
options(knitr.kable.NA = '--')


btwt <- spread(btwt,LIFESTAGE, BIOMASS)
#btwt <- select(btwt, "STOCKING_YEAR", "Fry", "Spring Fingerling", "Fall Fingerling", "Spring Yearling", "Adult")
btwt <- adorn_totals(btwt, where = "row")
btwt <- btwt %>% mutate(Total = rowSums(.[2:4], na.rm = TRUE))
btwt$Total <- round(btwt$Total,digits = 0)
btwt <- btwt %>%
  rename(
    YEAR = STOCKING_YEAR
  )

library(kableExtra)
kbl(btwt,format.args = list(big.mark = ","))%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = TRUE))%>%
  row_spec(nrow(btwt)-1, extra_css = "border-bottom: 1.5px solid")

```


<br><hr style="border:1px solid gray"> </hr><br>



## Stocking Dates
Average dates of Brown Trout stocking events (spring yearlings only)
```{r stocking_date}

# Check this carefully - surplus brood, etc., can be stocked at unusual times - omit these outliers.
dfdate <- df

#omit_stage <- c("Egg", "Adult", "Sub-adult", "Fry")
#dfdate <- subset(dfdate, !(LIFESTAGE %in% omit_stage))

dfdate <- filter(dfdate, dfdate$LIFESTAGE == "Spring Yearling")

dfdate$ORDINAL <- format(dfdate$STOCKING_EVENT_DATE, "%j")
dfdate$ORDINAL <- as.numeric(dfdate$ORDINAL)

releasedate <- tbl_df(dfdate) %>%
  group_by(STOCKING_YEAR, LIFESTAGE) %>%
  summarise(N = length(STOCKING_YEAR), avg = round(mean(ORDINAL, na.rm=T),2), 
            sd = round(sd(ORDINAL, na.rm=T),2), se = round(sd/sqrt(N),2)) 


# find average release date by life stage

stageavg <- releasedate %>%
  group_by(LIFESTAGE) %>%
  summarise(stageaverage = as.integer(mean(avg, na.rm=TRUE)))

releasedate <- left_join(releasedate, stageavg)

# # add these columns so a label can be built
releasedate$mon = 1
releasedate$day = 1
releasedate$start <- as.Date(with(releasedate, paste(STOCKING_YEAR, mon, day, sep = "-")), "%Y-%m-%d")
releasedate$average <- as.Date((releasedate$start) + releasedate$avg)
releasedate$datelabel <- format(releasedate$average, "%b %d")

# vertical plot

fig4 <-ggplot(releasedate, aes(x=STOCKING_YEAR, y=avg, label = datelabel)) + 
  ylab("Ordinal Day") + xlab("Year Stocked") +
  #labs (title = "Average Dates of Atlantic Salmon Stocking (life stage/year)")+ 
  geom_errorbar(aes(ymin=avg-sd, ymax=avg+sd), width=.5) +
  geom_point(colour="black", shape=19, size = 3) + 
  geom_hline(data = releasedate, aes(yintercept = stageaverage), color="grey", size = 1, linetype = 5) +
  theme(legend.title=element_blank())+scale_color_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous (breaks = seq(minyr, maxyr, by=2))+
  theme(legend.position = c(.2, .8)) + 
  theme_gfs()+ 
  theme(axis.text.x=element_text(size=16))+
  geom_text(hjust = 0, nudge_x = 0.12, size = 3) + 
  expand_limits(x = 2020.5)

fig4 + facet_grid(LIFESTAGE ~., scales="free") +
  theme(strip.background =element_rect(fill="white"))

```

<br><hr style="border:1px solid gray"> </hr><br>

## Stocking Temperatures

Average stocking site temperatures by year.  Overall average (all years) indicated by horizontal dashed line and red text annotation.

```{r Temperatures}
temp <- df

temp$SITE_TEMPERATURE <- as.numeric(temp$SITE_TEMPERATURE)


temp <- temp %>%
  group_by(STOCKING_YEAR, LIFESTAGE) %>%
  summarise(N = length(STOCKING_YEAR), avg = round(mean(SITE_TEMPERATURE, na.rm=T),2), 
            sd = round(sd(SITE_TEMPERATURE, na.rm=T),2), se = round(sd/sqrt(N),2)) 

temp <- filter(temp, LIFESTAGE == "Spring Yearling")

temp$rivavg <- round(temp$avg, digits = 1)

mean_temp <- round(mean(temp$avg),digits = 1)
 
 
temp$tlabel <- round(temp$avg, digits = 1)



fig5b <-ggplot(temp, aes(x=STOCKING_YEAR, y=avg, label = tlabel)) + 
  ylab("Temperature") + xlab("Year Stocked") +
  labs (title = "Spring Yearlings - avg. temp.")+ 
  geom_errorbar(aes(ymin=avg-sd, ymax=avg+sd), width=.5) +
  geom_point(colour="black", shape=19, size = 3) + 
  geom_hline(aes(yintercept = mean_temp), color="grey", size = 1, linetype = 5) +
  geom_text(aes(maxyr-1, mean_temp,label = mean_temp, vjust = -0.5), color = "red")+
  theme(legend.title=element_blank())+scale_color_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous (breaks = seq(minyr, maxyr, by=2))+
  theme(legend.position = c(.2, .8)) + 
  theme_gfs()+ 
  theme(axis.text.x=element_text(size=16))+
  #theme(axis.text.y=element_text(size=20))+
  geom_text(hjust = 0, nudge_x = 0.12, size = 3) + 
  expand_limits(x = maxyr + 0.5)

fig5b

# fig5b + facet_grid(WATERBODY ~., scales="free") +
#   theme(strip.background =element_rect(fill="white"))+
#   theme(panel.spacing = unit(2, "lines"))

```


<br><hr style="border:1px solid gray"> </hr><br>

## Map

Stocking events aggregated by stocking site (spring yearlings only).  Radius of circle indicates biomass (kg) stocked

```{r MAP}

zone <- readOGR("Data/Spatial/stockingzone.kml", verbose = FALSE)

# create stocking zone number labels
zonenum <- read.csv(textConnection("
Name,Lat,Long
1,43.316851,-79.281809
2,43.327948,-79.593005
3,43.497948,-79.407603
4,43.610478,-79.087773
5,43.77998,-78.407522
6,43.77998,-77.524854
7,43.77998,-76.872594
"))
zonenum$Name <- as.factor(zonenum$Name)

stocked <- df %>% subset(
  LIFESTAGE == "Spring Yearling"
)

stocked <- stocked%>%
  group_by(
    STOCKING_YEAR,
    STOCKING_ZONE,
    WATERBODY,
    LATITUDE,
    LONGITUDE,
    SITE)%>%
  summarise(
    NUMBER = sum(FISH_STOCKED_COUNT),
    BIOMASS = sum(BIOMASS_KG)
  )



palette <- colorFactor(palette = rainbow(length(unique(stocked$STOCKING_YEAR))), domain = stocked$STOCKING_YEAR)

sd <- SharedData$new(stocked)


map <- leaflet(width = "100%", height = 800) %>%
  addProviderTiles("Esri.WorldStreetMap", group = "Road map") %>%
  addProviderTiles("Esri.WorldImagery", group = "Satellite") %>%
  addProviderTiles("CartoDB.Positron", group = "Gray Scale") %>%
  addPolylines(data = zone, color='#2980B9', weight = 2, group = "Stocking Zones") %>%
  addCircleMarkers(data = sd, lat = ~LATITUDE, lng = ~LONGITUDE, radius = ~BIOMASS/50,
                   stroke = TRUE,
                   color = "black",
                   weight = 1,
                   fillColor = ~palette(stocked$STOCKING_YEAR),
                   fillOpacity = .5,
                   label=~as.character(SITE),
                   popup = paste
                   ("<b><font size=+1><font color=#5499C7>",stocked$SITE, "</b></font size></font color><br>",
                     "_______________________<br>",
                     "<br><b>Year: </b>", stocked$STOCKING_YEAR,
                     "<br><b>Number: </b>", stocked$NUMBER,
                     "<br><b>Biomass (kg): </b>", stocked$BIOMASS,
                     "<br><b>LAT: </b>", stocked$LATITUDE,
                     "<br><b>LON: </b>", stocked$LONGITUDE
                   ))%>%
   addLabelOnlyMarkers(data = zonenum, label = zonenum$Name ,group = 'Stocking Zones',
                      labelOptions = labelOptions(noHide = T, textOnly = TRUE,offset = c(0,1),
                                                  style=list(
                                                    'color'='#2980B9',
                                                    'font-family'= 'arial',
                                                    'font-style'= 'bold',
                                                    'font-size' = '25px'
                                                  )))%>%
  addLayersControl(position = 'topright',
                   baseGroups = c("Gray Scale",
                                  "Satellite",
                                  "Road Map"),
                   overlayGroups = c("Stocking Zones"))

bscols(widths = c(3,12),
       list(
         filter_select(id = "STOCKING_YEAR", label = "Year", sharedData = sd, group = ~STOCKING_YEAR, multiple = FALSE)
         #filter_slider(id = "STOCKING_YEAR", label = "Year", sharedData = sd, column = ~STOCKING_YEAR, step = 1, sep = "", width = "100%")
         ),
  map
)


```

<br><hr style="border:1px solid gray"> </hr><br>

<a href="#header">back to top</a>